A faster exact method for solving the robust multi-mode resource-constrained project scheduling problem

### 2022-03-13 paper/paper.tex
Responding to Marc's comments.

\newcommand{\marc}[1]{\textcolor{blue}{#1}}

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.9]{mrcpsp_network.pdf}
	\caption{An example instance involving five non-dummy activities and a single renewable resource. Activities 2 and 5 each have two available processing modes, whilst the other activities have only a single mode.}
	\label{fig:mrcpsp_network}
\end{figure}

@@

The aim of the two-stage robust MRCPSP is therefore to 
% select set of 
\marc{determine}
activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan. For the case where activity durations lie in a budgeted uncertainty set that we consider in this paper, this problem can be written as
\begin{align}
	\min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)}\ & \min S_{n+1}\label{eqn:robust_mrcpsp_1}\\
	\text{s.t. } & S_0 = 0\\
	& S_j - S_i \geq d_{im_i} \qquad \forall (i,j)\in E\cup X_{\bm{m}}\\
	& S_i \geq 0 \qquad \forall i\in V,\label{eqn:robust_mrcpsp_4}
\end{align}
where $\M\subseteq \mathbb{N}^n$ represents the set of all possible combinations of activity processing mode selections, and $\X_{\bm{m}}$ is the set of all possible sufficient selections for the choice of processing modes given by $\bm{m}$.

@@

This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network. This 
% augmented project network 
\marc{directed acyclic graph}
is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project \marc{(a similar construction has been applied in \cite{bendotti2019anchor})}. Having reformulated the adversarial subproblem as a longest-path problem, strong duality can be employed to enable its insertion it into a formulation for the first-stage problem, and ultimately arrive at a complete compact formulation for the full problem. Given that the derivation of this formulation is more or less identical to the derivation of the corresponding formulation for the robust RCPSP, we omit it here and instead refer the reader to \cite{bold2021compact}.

@@

The role of the master problem is to determine a choice of activity processing modes $\bm{m}$, as well as a sufficient selection $X_{\bm{m}}$ to resolve any resulting resource conflicts. This problem is solved without the direct consideration of the uncertain activity durations, and instead, uncertainty is accounted for in the subproblem and communicated back to the master problem with the use of optimality cuts. Hence, to remove the uncertainty from the model, (\ref{eqn:compact1})-(\ref{eqn:compact13}) is modified by considering only a single level of the augmented project network. Using this formulation, the master problem \marc{in iteration $t$} can be written as
\begin{align}
\min \ & \eta\label{eqn:master_first}\\
\text{s.t. } & \eta \geq S_{n+1}\\
	     & S_{0}=0\label{eqn:master2}\\
	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
	     & \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=1,\dots,t-1\label{eqn:master_cuts}\\
	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
	     & S_i\geq 0 \qquad \forall i\in V,\label{eqn:master_last}
\end{align}
where (\ref{eqn:master_cuts}) are the optimality cuts generated from the solutions of the resulting subproblems at each of the previous iterations.

@@

The instances used in this computational study have been created from the deterministic $j10$, and $j20$ MRCPSP instances from the PSPLIB (\cite{kolisch1997psplib}, \url{https://www.om-db.wi.tum.de/psplib/}). The $j10$ set contains a total of 536 instances each involving 10 activities, and the $j20$ set contains a total of 554 instances each involving 20 activities. We introduce uncertainty into these deterministic instances by setting the maximum durational deviation for each activity to be $\hat{d}_{im}=\lfloor0.7\times \bar{d}_{im}\rfloor$ for each mode $m\in M_i$. These uncertain instances have then been solved using both methods for a range of robustness levels $\Gamma$. In particular we solve the $j10$ instances for $\Gamma\in\{0,3,5,7\}$ and the $j20$ instances for $\Gamma\in\{0,5,10,15\}$.

@@

\begin{figure}[htb]
	\makebox[\textwidth][c]{
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j10_performance_profile.pdf}
		\caption{Performance profiles showing percentage of instances with a performance ratio of within $\tau$. Plotted on a log scale.}\label{fig:j10_performance_profile}
	\end{subfigure}%
	\qquad
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j10_gaps.pdf}
		\caption{Percentage of instances solved to within given optimality gap within the two-hour time limit.}\label{fig:j10_gaps}
	\end{subfigure}
	}
	\caption{Comparison of the Benders' approaches and compact formulation over instances in the $j10$ set.}
	\label{fig:j10_plots}
\end{figure}

\begin{figure}[htb]
	\makebox[\textwidth][c]{
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j20_performance_profile.pdf}
		\caption{Performance profile showing percentage of instances with a performance ratio of within $\tau$. Plotted on a log scale.}\label{fig:j20_performance_profile}
	\end{subfigure}%
	\qquad
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j20_gaps.pdf}
		\caption{Percentage of instances solved to within given optimality gap within the two-hour time limit.}\label{fig:j20_gaps}
	\end{subfigure}
	}
	\caption{Comparison of the Benders' approaches and compact formulation over instances in the $j20$ set.}
	\label{fig:j20_plots}
\end{figure}

@@

These performance profiles show that whenever an instance was able to be solved to optimality by either solution approach, the compact formulation was always the faster method. If we specifically compare the solution times of the compact formulation with the stronger of the two Benders' implementations, across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s). 
% And 
\marc{When} both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).

>>>>

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.9]{mrcpsp_network.pdf}
	\caption{An example instance involving five non-dummy activities and a single renewable resource with availability $R_1=4$. Activities 2 and 5 each have two available processing modes, whilst the other activities have only a single mode.}
	\label{fig:mrcpsp_network}
\end{figure}

@@

The aim of the two-stage robust MRCPSP is therefore to determine activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan. For the case where activity durations lie in a budgeted uncertainty set that we consider in this paper, this problem can be written as
\begin{align}
	\min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)}\ & \min S_{n+1}\label{eqn:robust_mrcpsp_1}\\
	\text{s.t. } & S_0 = 0\\
	& S_j - S_i \geq d_{im_i} \qquad \forall (i,j)\in E\cup X_{\bm{m}}\\
	& S_i \geq 0 \qquad \forall i\in V,\label{eqn:robust_mrcpsp_4}
\end{align}
where $\M\subseteq \mathbb{N}^n$ represents the set of all possible combinations of activity processing mode selections, and $\X_{\bm{m}}$ is the set of all possible sufficient selections for the choice of processing modes given by $\bm{m}$.

@@

This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network. This directed acyclic graph is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project (a similar construction has also been applied in \cite{bendotti2019anchor}). Having reformulated the adversarial subproblem as a longest-path problem, strong duality can be employed to enable its insertion it into a formulation for the first-stage problem, and ultimately arrive at a complete compact formulation for the full problem. Given that the derivation of this formulation is more or less identical to the derivation of the corresponding formulation for the robust RCPSP, we omit it here and instead refer the reader to \cite{bold2021compact}.

@@

The role of the master problem is to determine a choice of activity processing modes $\bm{m}$, as well as a sufficient selection $X_{\bm{m}}$ to resolve any resulting resource conflicts. This problem is solved without the direct consideration of the uncertain activity durations, and instead, uncertainty is accounted for in the subproblem and communicated back to the master problem with the use of optimality cuts. Hence, to remove the uncertainty from the model, (\ref{eqn:compact1})-(\ref{eqn:compact13}) is modified by considering only a single level of the augmented project network. Using this formulation, the master problem at iteration $t$ can be written as
\begin{align}
\min \ & \eta\label{eqn:master_first}\\
\text{s.t. } & \eta \geq S_{n+1}\\
	     & S_{0}=0\label{eqn:master2}\\
	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
	     & \eta \geq \lambda(\bm{x}, \bm{y}, V^{*\ell}) \qquad \forall \ell=1,\dots,t-1\label{eqn:master_cuts}\\
	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
	     & S_i\geq 0 \qquad \forall i\in V,\label{eqn:master_last}
\end{align}
where (\ref{eqn:master_cuts}) are the optimality cuts generated from the solutions of the resulting subproblems at each of the previous iterations.

@@

The instances used in this computational study have been created from the deterministic $j10$, and $j20$ MRCPSP instances from the PSPLIB (\cite{kolisch1997psplib}, \url{https://www.om-db.wi.tum.de/psplib/}). Note that these are the same instance sets as used for the experiments in \cite{balouka2021robust}. The $j10$ set contains a total of 536 instances each involving 10 activities, and the $j20$ set contains a total of 554 instances each involving 20 activities. We introduce uncertainty into these deterministic instances by setting the maximum durational deviation for each activity to be $\hat{d}_{im}=\lfloor0.7\times \bar{d}_{im}\rfloor$ for each mode $m\in M_i$. These uncertain instances have then been solved using both methods for a range of robustness levels $\Gamma$. In particular we solve the $j10$ instances for $\Gamma\in\{0,3,5,7\}$ and the $j20$ instances for $\Gamma\in\{0,5,10,15\}$.

@@

\begin{figure}[h]
	\makebox[\textwidth][c]{
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j10_performance_profile.pdf}
		\caption{Performance profiles showing percentage of instances with a performance ratio of within $\tau$. Plotted on a log scale.}\label{fig:j10_performance_profile}
	\end{subfigure}%
	\qquad
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j10_gaps.pdf}
		\caption{Percentage of instances solved to within given optimality gap within the two-hour time limit.}\label{fig:j10_gaps}
	\end{subfigure}
	}
	\caption{Comparison of the Benders' approaches and compact formulation over instances in the $j10$ set.}
	\label{fig:j10_plots}
\end{figure}

\begin{figure}[h]
	\makebox[\textwidth][c]{
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j20_performance_profile.pdf}
		\caption{Performance profile showing percentage of instances with a performance ratio of within $\tau$. Plotted on a log scale.}\label{fig:j20_performance_profile}
	\end{subfigure}%
	\qquad
	\begin{subfigure}[t]{.6\textwidth}
		\centering
		\includegraphics[width=1.05\textwidth]{j20_gaps.pdf}
		\caption{Percentage of instances solved to within given optimality gap within the two-hour time limit.}\label{fig:j20_gaps}
	\end{subfigure}
	}
	\caption{Comparison of the Benders' approaches and compact formulation over instances in the $j20$ set.}
	\label{fig:j20_plots}
\end{figure}

@@

These performance profiles show that whenever an instance was able to be solved to optimality by either solution approach, the compact formulation was always the faster method. If we specifically compare the solution times of the compact formulation with the stronger of the two Benders' implementations, across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s). When both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
